---
title: "Chapter 2"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Simple Linear Regression

<hr>
```{r setup, include=FALSE}
library(tidyverse)
```

```{r}
# Computer Repair Dataset

df <- read_delim("data/P031.txt", "\t", 
                 escape_double = FALSE, 
                 col_types = cols(Minutes = col_integer(), 
                                  Units = col_integer()), 
                 trim_ws = TRUE)

head(df)
```

Plot: 
```{r}
ggplot(df, aes(Units, Minutes)) + 
  geom_point() + 
  theme_minimal()
```

Covariance: 
```{r}
cov(df$Units, df$Minutes)
```

Correlation: 
```{r}
cor(df$Units, df$Minutes)
```

Formula for the coefficient $\hat\beta_1$:
$$\hat\beta_1 = \frac{Cov(X, Y)}{Var(X)}$$

```{r}
(covariance <- cov(df$Units, df$Minutes))
(variance_x <- var(df$Units))
(beta_hat_1 <- covariance / variance_x)
```

```{r}
table_2.7 <- df %>% 
  rename(x = Units, y = Minutes) %>% 
  mutate(y_hat = 4.162 + (15.509 * x), 
         e = y - y_hat)

table_2.7
```
```{r}
table_2.7 %>% 
  ggplot(aes(x, y)) + 
  geom_point() + 
  geom_line(aes(x, y_hat), size = 0.2) +
  theme_minimal()
```
```{r}
table_2.7 %>% 
  ggplot(aes(x, e)) + 
  geom_point() + 
  theme_minimal()
```
## Tests of hypotheses

```{r}
df2 <- table_2.7[,c(1, 2)] %>% 
  mutate(y_hat = 0 * x, 
         e = y - y_hat)

df2

df2 %>% 
  ggplot(aes(x, e)) + 
  geom_point() + 
  theme_minimal()
```
```{r}
model <- lm(Minutes ~ Units, data = df)

model

summary(model)
```
```{r}
15.509 / 0.505
```
```{r}
table_2.7 %>% 
  ggplot(aes(y, y_hat)) + 
  geom_point() + 
  theme_minimal()
```
```{r}
plot(model)
```

## Exercises

### 2.1

```{r}
# Var(Y) and Var(X)

var(df$Minutes)
var(df$Units)
```

Prove or verify that 
$$\sum_{i=1}^n(y_i - \bar y) = 0$$
```{r}
sum(df$Minutes - mean(df$Minutes))

# Comes out close to 0
```

### 2.3

```{r}
# a) 

# Test statistic: 
t1 <- (15.509 - 15) / 0.505

# Critical value:
t <- 2 * qt(0.05, 12)

t < t1 # Significant? No
```

### 2.10 Heights of husbands and wives

```{r}
df_2.10 <- read_delim("data/P052.txt", "\t", 
    escape_double = FALSE, trim_ws = TRUE)

df_2.10 <- df_2.10 %>% rename(X = Husband, Y = Wife)
```
```{r}
df_2.10 %>% 
  ggplot(aes(X, Y)) + 
  geom_point() + 
  theme_minimal() 
```
```{r}
# Reverse X and Y
df_2.10 %>% 
  ggplot(aes(Y, X)) +
  geom_point() +
  theme_minimal() 
```

```{r}
# See what it looks like if both scales start at 0

df_2.10 %>% 
  ggplot(aes(X, Y)) + 
  geom_point() + 
  theme_minimal() + 
  scale_x_continuous(limits = c(0, NA)) + 
  scale_y_continuous(limits = c(0, NA))
```

```{r}
# a) Covariance
cov(df_2.10$X, df_2.10$Y)
```
```{r}
# b) Covariance if heights are measured in inches

df_2.10 <- df_2.10 %>% 
  mutate(X_inch = X / 2.54, 
         Y_inch = Y / 2.54)

cov(df_2.10$X_inch, df_2.10$Y_inch)
```
```{r}
# c) Correlation CM
cor(df_2.10$X, df_2.10$Y)
```
```{r}
# d) Correlation Inches
cor(df_2.10$X_inch, df_2.10$Y_inch)
```
```{r}
# e) Correlation if every man married a woman exactly 5cm shorter than him
df_2.10 <- df_2.10 %>% 
  mutate(Y_5cm = X - 5)

cor(df_2.10$X, df_2.10$Y_5cm) # 1

```
```{r}
# f) Which should be response variable? 

#     Mathematically shouldn't make a difference. Comes down to social knowledge. Let's keep Y as response, wife's height 

```
```{r}
# g) Fit model, test null hypothesis that slope is 0. 

model_2.10 <- lm(Y ~ X, data = df_2.10)
summary(model_2.10)

```
```{r}
# Hypothesis that slope is 0

# Probability of t value
pt(-11.458, 94) * 2

# Result is significant (as shown above, minuscule p-value)
```

```{r}
# Model the other way

predict_husband <- lm(X ~ Y, data = df_2.10)
summary(predict_husband)
```

### 2.12 Newspaper circulation

```{r}
df_2.12 <- read_delim("data/P054.txt", "\t", 
    escape_double = FALSE, trim_ws = TRUE)
df_2.12
```
```{r}
# a) Scatterplot

df_2.12 %>% 
  ggplot(aes(Daily, Sunday)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(0, NA)) + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_minimal()
```
```{r}
# b) Fit model

model_news <- lm(Sunday ~ Daily, data = df_2.12)
summary(model_news)
```
```{r}
# c) Obtain 95% confint for the coefficients
#     User r confint, default level = 0.95

confint(model_news)
```
```{r}
# d) Is there justification? Yes, don't need to do full hypothesis test. R tells us p-value that slope != 0 is <2e-16. 

# e) What proportion of Sunday variability is accounted for by daily circulation? 

summary(model_news)$adj.r.squared # 91.5%
```
```{r}
# f) Provide an estimate (95% level) of the average Sunday circulation of newspapers with daily circulation of 500,000.

```
From page 41, we want an estimation of the mean reponse $\mu_0$ when $X = x_0$. This is:
$$ \hat\mu_0 = \hat\beta_0 + \hat\beta_1x_0$$
Finding the mean is the easy part, with R: 
```{r}
(mu_hat_0 <- summary(model_news)$coefficients[1] + summary(model_news)$coefficients[2] * 500)
```

Now we need confidence levels. Could do the formula manually, but R has it. See this article: https://rpubs.com/aaronsc32/regression-confidence-prediction-intervals

```{r}
new_dat <- data.frame(Daily = 500)
predict(model_news, newdata = new_dat, interval = "confidence")
```
```{r}
# Plot this

df_2.12 %>% 
  ggplot(aes(Daily, Sunday)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_x_continuous(limits = c(0, NA)) + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_minimal() + 
  geom_segment(aes(500, 644, xend = 500, yend = 723), color = "darkred") + 
  geom_point(aes(500, 684), size = 2, color = "darkred") + 
  labs(title = "Point estimate and confidence limits", 
       subtitle = "For mean Sunday if Daily = 500 (thousand)")
```
```{r}
# g) Predicted Sunday circulation for a given newspaper. Point estimate will be the same as above, but will use "prediction intervals" instead of "confidence intervals". 

new_dat <- data.frame(Daily = 500)
predict(model_news, newdata = new_dat, interval = "prediction")
```
```{r}
# Plot this

df_2.12 %>% 
  ggplot(aes(Daily, Sunday)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_x_continuous(limits = c(0, NA)) + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_minimal() + 
  geom_segment(aes(500, 457, xend = 500, yend = 910), color = "darkred") + 
  geom_point(aes(500, 684), size = 2, color = "darkred") + 
  labs(title = "Point estimate and prediction limits", 
       subtitle = "For Sunday circulation of given paper with Daily = 500 (thousand)")
```
```{r}
# h) Prediction for another paper with daily circulation of 2000 (i.e 2,000,000)

new_dat <- data.frame(Daily = 2000)
predict(model_news, newdata = new_dat, interval = "prediction")
```
```{r}
# Plot this

df_2.12 %>% 
  ggplot(aes(Daily, Sunday)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_x_continuous(limits = c(0, NA)) + 
  scale_y_continuous(limits = c(0, NA)) +
  theme_minimal() + 
  geom_segment(aes(2000, 2373, xend = 2000, yend = 3013), color = "darkred") + 
  geom_point(aes(2000, 2693), size = 2, color = "darkred") + 
  labs(title = "Point estimate and prediction limits: out of sample", 
       subtitle = "For Sunday circulation of given paper with Daily = 2000 (thousand)")
```
